Evaluating Cultural Transmission in Bronze Age burial rites of Central, Northern and North-western Europe using radiocarbon data

### 2019-06-29 article/05_Conclusion.Rmd
proofreading

From the point of view of Cultural complexity -- the question, which factors accelerate or decelerate the variability of human behaviour [@henrichDemographyCulturalEvolution2004; @mesoudiWhatCumulativeCultural2018; @vaesenPopulationSizeDoes2016] -- the long term dominance fluctuations of one burial rite over the other are remarkable. Its standard explanation models (among others population density and environmental risk) can hardly be applied to stylistic variables.

>>>>

From the point of view of *Cultural complexity* -- the question, which factors accelerate or decelerate the variability of human behaviour [@henrichDemographyCulturalEvolution2004; @mesoudiWhatCumulativeCultural2018; @vaesenPopulationSizeDoes2016] -- the long term dominance fluctuations of one burial rite over the other are remarkable. Its standard explanation models (among others population density and environmental risk) can hardly be applied to stylistic variables.

### 2019-06-29 article/06_Supplementary_Remarks.Rmd
proofreading

5. In very few regions and periods of the European Bronze Age the number of excavated graves is sufficiently big to assume that a statistically relevant sample of the burial culture of the entire population is documented. Instead, it is evident in most contexts that only the burials of certain social groups, genders or ethnic groups were recorded. The quantity and nature of the known graves cannot be representative. In this case it can be expected that further, deviant burial rituals were practised, which left no archaeologically visible remains. These rituals are invisible in the ^14^C data from Radon-B.

@@

The superordinate, large perspective also tries to simplify and categorize the parameters associated with a single burial. Radon-B anticipated this step by providing the variables *burial type* and *burial construction* that can be reduced to three categories each. This is extremely useful on the one hand, but also questionable on the other. The respective contexts can be multifarious: similar rituals in different cultural contexts can have different meanings. For this paper, it is assumed -- with reference to *Cultural Evolution Theory* -- that despite this restriction, actions such as burning the dead or burying in burial mounds can be passed on across cultural and ethnic group boundaries through imitation and teaching. To put it provocatively: cremation burial in the Hungarian Early Bronze Age is a manifestation of the same idea as cremation in the Northern Late Bronze Age. This is a working hypothesis, but may be questioned.

>>>>

5. In very few regions and periods of the European Bronze Age the number of excavated graves is sufficiently big to assume that a statistically relevant sample of the burial culture of the entire population is documented. Instead, it is evident in most contexts that only the burials of certain social groups, genders or ethnic groups were recorded. The quantity and nature of the known graves cannot be representative. In this case it can be expected that other burial rituals were practised, which left no archaeologically visible remains. These rituals are invisible in the ^14^C data from Radon-B.

@@

The superordinate, large perspective also tries to simplify and categorize the parameters associated with a single burial. Radon-B anticipated this step by providing the variables *burial type* and *burial construction* that can be reduced to three categories each. This is extremely useful on the one hand, but also questionable on the other. The respective contexts can be multifarious: similar rituals in different cultural contexts can have different meanings. For this paper, it is assumed -- with reference to *Cultural Evolution Theory* -- that despite this restriction, behaviour such as burning the dead or burying in burial mounds can be passed on across cultural and ethnic group boundaries through imitation and teaching. To put it provocatively: cremation burial in the Hungarian Early Bronze Age is a manifestation of the same idea as cremation in the Northern Late Bronze Age. This is a working hypothesis, but may be questioned.
